Algebraic Principles for Rely-Guarantee Style Concurrency Verification Tools

### 2013-11-11 Paper/paper.tex
Started on introduction

\title{Algebraic Principles for Concurrency Verification Tools}

\author{Alasdair Armstrong \and Victor B.~F.~Gomes\thanks{Sponsored by CNPq, Brazil}  \and Georg Struth}

@@

\begin{abstract}
  We investigate equational laws for the derivation of rely-guarantee
  style inference and refinement rules in the minimalistic setting of
  semirings. This algebraic level is linked with concrete programming
  models based on languages and execution traces. The approach has
  been implemented in Isabelle/HOL where the link between algebras and
  models supports reasoning about both the control and data flow of
  concurrent programs at varying levels of abstraction. Finally, we
  illustrate our approach via simple program verification examples.
\end{abstract}

@@

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

x

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Semirings and Kleene Algebra}
\label{sec:KA}

A \emph{semiring} is an structure
$(D,+,\cdot,0,1)$ satisfying the following laws,
\begin{align*}
  x(y + z) &= xy + xz, &\qquad (x + y)z &= xz + yz,\\
  (x + y) + z &= x + (y + z), &\qquad (xy)z &= x(yz),\\
  x + y &= y + x, &\qquad 0 + x &= x,\\
  1x &= x, &\qquad x1 &= x,\\
  0x &= 0, &\qquad x0 &= 0.
\end{align*}
A semiring is \emph{idempotent} if $x + x = x$, and \emph{weak} if $x0
= 0$ does not hold. The natural partial order $\le$ is defined as $x
\le y $ iff $x + y = y$. In an idempotent semiring $(D,+,0)$ therefore
forms a join semilattice with least element $0$. Idempotent semirings
are also called \emph{dioids}. In some contexts, it is also useful to
add a meet operation $\sqcap$, such that $(D,+,\sqcap)$ is a lattice.

In the context of programs, one typically thinks of $\cdot$ as being
sequential composition, $+$ as non-deterministic choice, $0$ as the
abortive action and $1$ as skip. For concurrent software, it is
natural to extend the notion of a dioid to that of a \emph{trioid}
with operators for both sequential and parallel composition. Formally,
a trioid is a structure $(D,+,\cdot,\|,0,1)$ where $(D,+,\cdot,0,1)$
is a dioid and,
\begin{align*}
  x\|(y + z) &= x\|y + x\|z, &\qquad x\|y &= y\|x,\\
  (x\|y)\|z &= x\|(y\|z), &\qquad x\|1 &= x,\\
  0\|x &= x.
\end{align*}

In a trioid there are no laws governing the interaction between
sequential and parallel composition. One law which has been considered
in previous papers is the \emph{exchange} law~\cite{hoare_concurrent_2011,hoare_locality_2011}
\begin{align*}
  (w \| x) \cdot (y \| z) \le (w \cdot z) \| (x \cdot y).
\end{align*}

The exchange law is initially appealing as it provides a way to derive
many properties of parallel composition from a single law. For example
one can show that $\|$ must be commutative, or that $x\cdot y \le
x\|y$. However, there are many interesting models which do not satisfy
this law. First, one could consider fair parallel composition
$x\parallel_f y$ in a model with infinite, or non-terminating
programs. In this model, $x \cdot y \not\leq x \parallel_f y$ in the
case where $x$ is non-terminating. A second example would be
predicate-transformer style models for separation logic, where $\|$ is
interpreted as separating conjunction. In such models, $\|$ and
$\cdot$ have different units, which, in the presence of the exchange
law, implies that $\cdot$ is equal to $\|$.

A Kleene algebra~\cite{kozen_completeness_1994} expands a dioid with a star operation satisfying the laws
\begin{align*}
  1 + xx^* &\le x^*, &\qquad z + xy \le y &\implies x^*z \le y,\\
  1 + x^*x &\le x^*, &\qquad z + yx \le y &\implies zx^* \le y.
\end{align*}

For programs, the star represents finite iteration of a program $x$
zero or more times. For iteration where $x$ must be executed at least
once, we write $x^+$. One can similarly axiomatise a parallel star
operation.

A Kleene algebra expanded with a parallel composition operator such
that $(K,+,\cdot,\|,0,1)$ is a trioid in which the exchange law holds
is called a \emph{concurrent Kleene algebra}~\cite{hoare_concurrent_2011}.

In a (concurrent) Kleene algebra Hoare triples can be encoded \`a la
Tarlecki as $\{x\}y\{z\} \iff xy \le z$. Assume residuals axiomatised
by Galois connections
\begin{align*}
 x \le z \leftarrow y \iff xy \le z \iff y \le x \rightarrow z,
\end{align*}
exist, as in Pratt's \emph{action
  algebra}~\cite{pratt_action_1990}. Then the Hoare triple
$\{x\}y\{z\}$ is equivalent to $x \le z \leftarrow y$, where $z
\leftarrow y$ can be seen as $\wlp(y,z)$.

For sequential programs, Hoare logic can also be encoded algebraically by
expanding a Kleene algebra with a Boolean test subalgebra. A
\emph{Kleene algebra with tests} (KAT)~\cite{kozen_kleene_1997} is structure
$(K,B,+,\cdot,\phantom{|}^\star,\overline{\phantom{x}},0,1)$ where
$(K,+,\cdot,\phantom{|}^\star,0,1)$ is a Kleene algebra and
$(B,+,\cdot,\overline{\phantom{x}},0,1)$ is a Boolean algebra such
that $B \subseteq K$. The $\overline{\phantom{x}}$ operator is only
defined on elements of the Boolean subalgebra. We denote arbitrary
elements of $K$ by $x,y,z$ and tests by $p,q,r$.

In KAT Hoare triples are defined in the following (equivalent) ways
\begin{align*}
\{p\}x\{q\} \iff px \le xq \iff px = pxq \iff px\overline{q} = 0.
\end{align*}
Using these encodings all the rules of Hoare logic excluding the
assignment rule can be derived.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebra for Rely-Guarantee}
\label{sec:RG}

A rely-guarantee algebra is a structure
$(K,RG,+,\sqcap,\cdot,\|,^\star,0,1)$, where $(K,+,\sqcap)$ is a
distributive lattice, $(K,+,\cdot,\|,0,1)$ is a weak trioid and
$(K,+,\cdot,^\star,0,1)$ is a Kleene algebra. $RG$ is a distinguished
subset of relys and guarantees which satisfy the following axioms
\begin{align}
r\|r &\le r, \label{rg1}\\
r &\le r\|r', \label{rg2}\\
r\|xy &= (r\|x)(r\|y), \label{rg3}\\
r\|x^+ &\le (r\|x)^+ \label{rg4}.
\end{align}
By convention, we use $r$ and $g$ to refer to elements of $RG$, and
$x,y,z$ for arbitrary elements of $K$. Some elementary consequences of these rules are as follows
\begin{align*}
1 &\le r,\\
r^\star = rr &= r = r\|r,\\
r\|x^+ &= (r\|x)^+.
\end{align*}

@@

If we expand our rely-guarantee algebras to have residuals axiomatised
by Galois connections for both parallel and sequential composition,
i.e.
\begin{align}
xy \le z &\iff y \le x \rightarrow z, \label{res1}\\
xy \le z &\iff x \le z \leftarrow y, \label{res2}\\
x\|y \le z &\iff y \le x/z \label{res3},
\end{align}
then for sequential composition, (\ref{res1}) and (\ref{res2}) provide
a link between weakest-precondition style semantics and hoare
logic. As mentioned in Section \ref{sec:KA}, the triple $\{p\} c
\{q\}$ can be encoded as $pc \le q$, while $q \leftarrow c$ represents
$wlp(c,q)$. Equation (\ref{res2}) can thus be written as $\{p\} c
\{q\} \iff p \le wlp(c,q)$. The parallel residual $x/z$ can be
understood similarly as follows; it is the weakest program such that
when placed in parallel with $x$, the composition behaves as $z$.

@@

>>>>

\title{Algebraic Principles for Rely-Guarantee Style Concurrency Verification Tools}

\author{Alasdair Armstrong \and Victor B.~F.~Gomes \and Georg Struth}

@@

\begin{abstract}
  We provide simple equational principles for deriving
  rely-guarantee-style inference rules and refinement laws based on
  idempotent semirings. We link the algebraic layer with concrete
  models of programs based on languages and execution traces. We have
  implemented the approach in Isabelle/HOL as a lightweight
  concurrency verification tool that supports reasoning about the
  control and data flow of concurrent programs with shared variables
  at different levels of abstraction. This is illustrated on two
  simple verification examples.
\end{abstract}

@@

We aim to develop tools which support reasoning about concurrent
programs at varying levels of abstraction. At the highest level,
abstract algebra provides a framework for the prototyping of new
inference and refinement rules, whereas at the lowest level concrete
models of program stores allow for fine-grained reasoning about
program data flow. We have implemented this approach is the
Isabelle/HOL interactive theorem prover. Isabelle/HOL not only
provides us with a high degree of confidence in the correctness of our
theorems, but also supports the construction of custom proof tactics
and tools for program verification tasks. Our approach differs from
existing frameworks for rely-guarantee reasoning in
Isabelle~\cite{nieto}, both by making the link between concrete
programs and algebras explicit, but also by allowing arbitrary nested
parallelism.

For sequential programs the applicability of algebra, and Kleene
algebra in particular, has been known for over a decade. Kleene
algebra provides operations for non-deterministic choice, sequential
composition and finite iteration, in addition to skip and abort. With
appropriate extensions, Kleene algebras support Hoare-style program
verification, and allow the derivation of program equivalences and
refinement rules. Furthermore, Kleene algebra has been applied in many
applications, such as compiler
optimisation~\cite{kozen_certification_2000}, program construction,
transformation and termination analysis, and static analysis.

Algebra has also been applied in the study of separation logic, for
example, as the separation algebras in~\cite{ohearn}, or in algebraic
separation logic.

\fbox{Algebraic Separation Logic}

\fbox{Concurrent Kleene Algebra}

Rely-guarantee is a so called \emph{compositional} method, where the
verification of a large program is reduced the independent
verification of individual subprograms, with the interference from
other subprograms being abstractly captured by \emph{rely} and
\emph{guarantee} conditions. This form of compositionality is,
however, weaker than the strong notion of compositionality inherent in
algebra, where full congruence rules exist for every operator.

In this paper, we investigate algebraic principles for rely-guarantee
style reasoning. This has been explored previously, for example
in~\cite{cka}, but we extend this approach by considering a minimal
set of rules sufficient to derive the standard rely-guarantee
inference rules. Furthermore, we explore how the algebraic
compositionality of this approach can be broken in just the right way,
so as to capture the ideas of abstract interference conditions in
rely-guarantee reasoning.

The verification and construction of actual concurrent programs often
requires complex semantics involving synchronisation, termination
analysis and complicated data structures.

\fbox{We verify simple programs}

\fbox{Coherent framework for programs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebraic Preliminaries}
\label{sec:KA}

\fbox{define weak trioid!}

Rely-guarantee algebras, which are introduced in the following
section, are based on dioids and Kleene algebras. A \emph{semiring} is
a structure $(S,+,\cdot,0,1)$ such that $(S,+,0)$ is a commutative
monoid, $(S,\cdot, 1)$ is a monoid and the distributivity laws $x\cdot
(y+z)=x\cdot z + y \cdot z$ and $(x+y)\cdot z = x\cdot z+y\cdot z$ as
well as the annihilation laws $x\cdot 0=0$ and $0\cdot x=0$ hold. A
\emph{dioid} is a semiring in which addition is idempotent:
$x+x=x$. Hence $(S,+,0)$ forms a join semilattice with least element
$0$ and partial order defined, as usual, as $x\le y\Leftrightarrow
x+y=y$. The operations of addition and multiplication are isotone with
respect to the order, that is, $x \le y $ implies $z+x\le z+y$,
$z\cdot x \le z\cdot y$ and $x\cdot z \le y\cdot z$. A dioid is
\emph{commutative} if multiplication is: $x\cdot y = y \cdot x$.

In the context of sequential programs, one typically thinks of $\cdot$
as sequential composition, $+$ as nondeterministic choice, $0$ as the
abortive action and $1$ as skip.  In this context it is essential that
multiplication is not commutative.  More formally, it is well known
that (regular) languages with language union as $+$ and language
product as $\cdot$, the empty language as $0$ and the empty word
language $\{\varepsilon\}$ as $1$ form dioids. Another model is formed
by binary relations with the union of relations as $+$, the product of
relations as $\cdot$, the empty relation as $0$ and the identity
relation as $1$. A model of commutative dioids is formed by sets of
(finite) multisets or Parikh vectors with multiset addition as
multiplication.

It is well known that commutative dioids can be used for modelling the
interaction between concurrent composition and nondeterministic
choice. The following definition serves as a basis for models of
concurrency in which sequential and concurrent composition
interact.\fbox{should we write $;$ instead of $\cdot$?}

A \emph{trioid} is a structure $(S,+,\cdot,||,0,1)$ such that
$(S,+,\cdot,0,1)$ is a dioid and $(S,\cdot,||,0,1)$ a commutative
dioid. In a trioid there is no interaction between the sequential
composition $\cdot$ and the parallel composition $||$. On the one
hand, Gischer has shown that trioids are sound and complete for the
equational theory of series-parallel pomset languages~\cite{Gischer},
which form a well studied model of true concurrency. On the other
hand, he has also obtained a completeness result with respect to a
notion of pomset subsumption for trioids with the additional
\emph{interchange axiom} $(w \| x) \cdot (y \| z) \le (w \cdot y) \|
(x \cdot z)$ and it is well known that this additional axiom also
holds for (regular) languages in which $||$ is interpreted as the
shuffle or interleaving operation~\cite{Gischer}.

Formally, the \emph{shuffle} $\sha$ of two finite words is defined
inductively as
\begin{equation*}
\epsilon \sha s = \{s\}, \qquad s \sha \epsilon = \{s\},\qquad
as \sha bt = a(s \sha bt) \cup b(as \sha t),
\end{equation*}
which is then lifted to the shuffle product of languages $X$ and $Y$
as
\begin{equation*}
X \sha Y = \{x \sha y:\; x \in X \land x \in Y\}.
\end{equation*}
\fbox{I would use $||$ here}

For logics of programs, notions of iteration or recursion are
essential. A \emph{Kleene algebra} is a dioid expaned with a star
operation which satisfies the \emph{left unfold axiom} $1 + x\cdot
x^*\le x^*$ and the \emph{left} and \emph{right induction axiom} $z+x
\cdot y \le y\Rightarrow x^*\cdot z \le y$ and $z + y\cdot x \le y
z\cdot \Rightarrow x^* \le y$. It follows that $1+x\cdot x^* = x^*$
and that the right unfold axiom $1+x^*\cdot x=x^*$ is derivable as
well. Thus iteration $x^\ast$ is modelled as the least fixpoint of the
function $\lambda y. 1+x\cdot y$, which is the same as the least
fixpoint of $\lambda y.1+ y\cdot x$. A \emph{commutative Kleene
  algebra} is a Kleene algebra in which multiplication is commutative.

It is well known that (regular) languages form Kleene algebras and
that (regular) sets of multisets form commutative Kleene algebras. In
fact, Kleene algebras are complete with respect to the equational
theory of regular languages as well as the equational theory of binary
relations with the reflexive transitive closure operation as the
star~\cite{Kozen}. Moreover, commutative Kleene algebras are
complete with respect to the equational theory of regular languages
over multisets~\cite{Conway}. It follows that equations in
(commutative) Kleene algebras are decidable.

A \emph{bi-Kleene algebra} is a structure
$(K,+,\cdot,||,0,1,^*,^{(*)})$ such that $(K,+,\cdot,0,1,^*)$ is a
Kleene algebra and $(K,+,||,0,1,^{(*)})$ a commutative Kleene
algebra. Bi-Kleene algebras are sound and complete with respect to the
equational theory of regular series-parallel pomset languages, and the
equational theory is again decidable~\cite{LaurenceStruth}. A
\emph{concurrent Kleene algebra} is a bi-Kleene algebra which
satisfies the interchange law~\cite{Hoareetal}. It can be shown that
shuffle languages and regular series-parallel pomset languages with a
suitable notion of pomset subsumption form concurrent Kleene algebras.

In some contexts, it is also useful to add a meet operation $\sqcap$,
such that $(D,+,\sqcap)$ is a lattice. \fbox{expand}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Generalised Hoare Logics  in Kleene Algebra}

It is well known that the inference rules for sequential Hoare logic
(except the assignment axiom) can be derived in expansions of Kleene
algebras. One approach is as follows~\cite{MoellerStruth}. Suppose a
suitable Boolean algebra $B$ of \emph{tests} has been embedded into a
Kleene algebra $K$ such that $0$ and $1$ are the minimal and maximal
element of $B$, $+$ corresponds to join and $\cdot$ to
meet. Complements $-$ are defined only on $B$. Suppose further that a
\emph{backward diamond operator} $\langle x|p$ has been defined for
each $x\in K$ and $p\in B$, which models the set of all states to
which each terminating execution of program $x$ may lead from states
$q$. Finally suppose that a \emph{forward box operator} $|x]p$ has
been defined which models the (largest) set of states from which every
terminating execution of $x$ must end up in states $q$ and that boxes
and diamonds are adjoints of the Galois connection
\begin{equation*}
  \langle x|p \le q \Leftrightarrow p \le |x]q,
\end{equation*}
for all $x\in K$ and $p,q\in B$. It is then evident from the above
explanations that validity of a Hoare triple $\vdash \{p\}x\{q\}$ can
be encoded as $\langle x|p \le q$ and the weakest liberal precodition
operator $\mathsf{wlp}(x,q)$ as $|x]p$. Hence the relationship between
the proof theory and the semantics of Hoare logic is captured by the
Galois connection $ \vdash\{p\}x\{q\} \Leftrightarrow p \le
\mathsf{wlp}(x,q)$. It has been shown that the relational semantics of
sequential while-programs can be encoded in these \emph{modal Kleene
  algebras} and that the inference rules of Hoare logic can be
derived~\cite{MoellerStruth}.

In the context of concurrency, this relational approach is no longer
appropriate. Here, the following approach by
Tarlecki~\cite{tarlecki_language_1985} can be used instead. One can
now encode validity of a Hoare triple as
\begin{equation*}
  \vdash \{x\}y\{z\} \Leftrightarrow x \cdot y \le z
\end{equation*}
for arbitrary elements of a Kleene algebra $K$. Nevertheless all the
rules of sequential Hoare logic except the assignment axiom can still
be derived in this setting~\cite{Hoareetal}.  Tarlecki's motivating
explanations carry over to the algebraic approach.

As an example we show the derivation of a generalised while
rule. Suppose $x\cdot t\cdot y\le x$. Then $x\cdot (t\cdot y)^\ast \le
x$ by the right induction axiom of Kleene algebra and therefore
$x\cdot (t\cdot y)^\ast t'\le x\cdot t'$ for arbitrary element $t'$ by
isotonicity of multiplication. This derives the while rule
\begin{equation*}
  \frac{\{x\cdot t\}y\{x\}}{\{x\}(t\cdot y)^\ast\{t'\cdot x\}}
\end{equation*}
for a generalised while loop $(t\cdot y)^\ast\cdot t'$, which
specialises to the conventional rule when $t$ and $t'$ are, in some
sense, complements.

The correspondence to a wlp-style semantics, as in model Kleene
algebra, now requires a generalisation of the Galois connection for
boxes and diamonds to multiplication and an upper adjoint in the form
of residuation.  This can be achieved in the context of \emph{action
  algebras}~\cite{pratt}, which expand Kleene algebras by operations
of left and right residuation defined by the Galois connections
\begin{equation*}
  xy\le z\Leftrightarrow x\le z\leftarrow y,\qquad xy\le z \Leftrightarrow y\le z\rightarrow x.
\end{equation*}
\fbox{please check that I didn't confuse these arrows} These
residuals, and now even the Kleene star can be axiomatised
equationally in action algebras. For a comprehensive list of the
properties of action algebras and their most important models
see~\cite{archive}, including the language and the relational
model. In analogy to the development in model Kleene algebra we can
now stipulate $ \wlp(x,y)=y\leftarrow x$ and obtain the Galois
connection
\begin{equation*}
  \vdash \{x\}y\{z\} \Leftrightarrow x\le \wlp(y,z)
\end{equation*}
with the characteristic properties $\vdash \{\wlp(y,z)\}y\{z\}$ and $x\le
\wlp(y,z)\Rightarrow \ \vdash \{x\}y\{z\}$. Moreover, if the action algebra is
also a quantale, and infinite sums exist, it follows that $
\wlp(y,z)=\sum\{x:\ \vdash \{x\}y\{z\}\}$.  It is obvious that this definition
makes sense in all models of action algebras and
quantales. Intuitively, suppose $p$ stands for the set of all
behaviours of a system, for instance the set of all execution traces,
that end in state $p$, and likewise for $q$. Then $\{p\}x\{q\}$ states
that all executions ending in $p$ can be extended by $x$ to executions
ending in $q$. $\wlp(x,q)$ is the most general behaviour, that is the
set of all executions $p$ after which all executions of $x$ must end
in $q$.

In the context of concurrent Kleene algebra, even without the parallel
star, additional inference rules can be derived, in particular 
\begin{equation*}
  \frac{\vdash\{x\}y\{z\}}{\vdash\{w||x\}y\{w||z\}}\qquad \frac{\vdash\{x_1\}y_1\{z_1\}\quad\vdash\{x_2\}y_2\{z_2\}}{\vdash\{x_1||x_2\}y_1||y_2\{z_1||z_2\}}
\end{equation*}
which correspond to the frame rule and the concurrency rule of
concurrent separation logic~\cite{} when $||$ is interpreted as
separating conjunction. In fact, the second inference rule is, in any
bi-Kleene algebra equivalent to the interchange axiom and the frame
rule is derivable from this axiom as well~\cite{Hoareetal}. The
precise connection between concurrent separation logic, a variant of
concurrent Kleene algebra and a resource transformer model has been
elaborated in~\cite{localitypaper}.

A residuation for concurrent composition can be considered as well:
\begin{equation*}
x\|y \le z \Leftrightarrow y \le x/z.
\end{equation*}
The residual $x/z$ represents the weakest program such that when
placed in parallel with $x$, the parallel composition behaves as $z$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A  Rely-Guarantee Algebra}
\label{sec:RG}

In this section we show how bi-Kleene algebras can be expanded into a
simple algebra that supports the derivation of rely-guarantee-style
inference rules. This development does \emph{not} use the interchange
law for several reasons. First, this law fails for fair parallel
composition $x\parallel_f y$ in a model with possibly infinite, or
non-terminating programs. In this model, $x \cdot y \not\leq
x \parallel_f y$ in the case whenever $x$ is non-terminating. Second,
it is not needed for deriving the usual concurrency laws. Third,
whereas it may be assumed in the models outlined in Section *** it
fails for the Aczel trace model in Section ***. \fbox{is that true?}

\fbox{this needs to be reworked. two letters for a set are not good}

\fbox{ the set RG doesn't need to be named in the algebra I think.}

\fbox{we can just start with a bi-KA and say that elements are relies/guarantees if they satisfy properties}

\fbox{what is a weak trioid?}

\fbox{say that we don't consider the parallel star}

\fbox{should we call relies and guarantees more generically
  ``conditions'' can call them $c$?}

A rely-guarantee algebra is a structure
$(K,RG,+,\sqcap,\cdot,\|,^\star,0,1)$, where $(K,+,\sqcap)$ is a
distributive lattice, $(K,+,\cdot,\|,0,1)$ is a weak trioid and
$(K,+,\cdot,^\star,0,1)$ is a Kleene algebra. $RG$ is a distinguished
subset of relys and guarantees which satisfy the following axioms
\begin{align}
r\|r &\le r, \label{rg1}\\
r &\le r\|r', \label{rg2}\\
r\|xy &= (r\|x)(r\|y), \label{rg3}\\
r\|x^+ &\le (r\|x)^+ \label{rg4}.
\end{align}
By convention, we use $r$ and $g$ to refer to elements of $RG$, and
$x,y,z$ for arbitrary elements of $K$.

Some elementary consequences of these rules are as follows
\begin{align*}
1 &\le r,\\
r^\star = rr &= r = r\|r,\\
r\|x^+ &= (r\|x)^+.
\end{align*}

@@

@@

CNPq EPSRC
